Physics of Tidal Dissipation in Early-Type Stars and White Dwarfs: Hydrodynamical Simulations of Internal Gravity Wave Breaking in Stellar Envelopes

### 2020-02-20 writeups/Su_IGW_break.tex
better movie plotting

% TODO
% - Figure 1: Could you include a tick mark at t=0? Does the curve start at t=0?
% Section 6.2 needs more narrative discussion of the results. We should add another paragraph saying that the typical Richardson number of the critical layer is between 0.25 and 0.5, characteristic of Kelvin-Helmholtz instabilities. This suggestions the critical layer width is self-consistently determined by the ``turbulence’’ in our simulations. The layer width does not change with resolution. We should also say that here we calculate Ri using the background buoyancy frequency, because the fluctuations do not change N^2 significantly; thus, Ri~0.5 can be used to solve for the critical layer width in a star, as we know N^2 and U_c (corresponding to the tidal frequency Omega).
% rewrite eq_fadef in explicit integrals
% Investigate a smarter time averaging for the "average over everything
% predictor"?
    \documentclass[
        fleqn,
        usenatbib,
        % referee,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        ae, aecompl,
        graphicx,
        booktabs,
    }
    \usepackage[T1]{fontenc}
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[caption=false]{subfig} % subfigures

@@

The geometry of our simulation domain is fixed by one further parameter, $z_0$,
the forcing location. We choose $L_z = 12.5H$, and the lower and upper damping
zones are located at $z < 1.5H$ and $z > 10.5H$ respectively. The forcing (see
Eq.~\eqref{eq:vol_drive}) is at $z_0 = 3H$ with width $\sigma = 0.078H$,
sufficiently far from the lower damping zone and permitting sufficient room for
the upward propagating wave to grow as $\propto e^{z/2H}$. Finally, we want
similar grid spacing in the $x$ and $z$ directions (i.e.\ $L_x / N_x \sim L_z
/N_z$), guided by the intuition that turbulence generated by wave breaking is
approximately isotropic, so we use $L_x = 4H$ and $N_z / N_x = 4$.

@@

After measuring $z_c$ (see Section~\ref{ss:cl_prop}) and $F(z)$
(Eq.~\eqref{eq:F_def}) at each time step, we determine each of $F_i$, $F_a$,
$F_r$, $F_s$ as follows:
\begin{align}
    F_i(t) &= F_{al}A_i^2(t),\\
    F_r(t) &= F_i(t) - \ev{F(z, t)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
        ,\label{eq:fr_def}\\
    F_s(t) &= \ev{\z{F(z, t): F(z, t) < 0}}_{z \in [z_c, z_c + \Delta z]},
        \label{eq:fs_def}\\
    F_a(t) &= F_i - F_r - F_s.\label{eq:fa_def}
\end{align}
Fig.~\ref{fig:nl_f_amps2} depicts the four components of this flux
decomposition.
% Here, $\ev{\dots}_{z \in [z_a, z_b]}$ denotes a vertical average over the
% interval $[z_a, z_b]$.
Below the critical layer, we average over an interval of length $H$, also the
vertical wavelength. The offset $\Delta z$ is necessary to make the measurement
of the incident flux unaffected by the turbulence within the critical layer
itself. The width of the critical layer is limited by $\mathrm{Ri} \lesssim 1$
(see Section~\ref{ss:khi}), which bounds its vertical extent $\sim
\frac{1}{\abs{k_{z}}}$. We empirically found an offset of $\Delta z =
\frac{3}{\abs{k_z}}$ was necessary to be sufficiently far from strong
fluctuations near the critical layer.

>>>>

% TODO
% - Figure 1: Could you include a tick mark at t=0? Does the curve start at t=0?
% Section 6.2 needs more narrative discussion of the results. We should add another paragraph saying that the typical Richardson number of the critical layer is between 0.25 and 0.5, characteristic of Kelvin-Helmholtz instabilities. This suggestions the critical layer width is self-consistently determined by the ``turbulence’’ in our simulations. The layer width does not change with resolution. We should also say that here we calculate Ri using the background buoyancy frequency, because the fluctuations do not change N^2 significantly; thus, Ri~0.5 can be used to solve for the critical layer width in a star, as we know N^2 and U_c (corresponding to the tidal frequency Omega).
% Investigate a smarter time averaging for the "average over everything
% predictor"?
    \documentclass[
        fleqn,
        usenatbib,
        % referee,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        ae, aecompl,
        graphicx,
        booktabs,
    }
    \usepackage[T1]{fontenc}
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[caption=false]{subfig} % subfigures

@@

The geometry of our simulation domain is fixed by one further parameter, $z_0$,
the forcing location. We choose $L_z = 12.5H$, and the lower and upper damping
zones are located at $z < 0.3H$ and $z > 9.5H$ respectively ($z \in [-1,
11.5]$). The forcing (see Eq.~\eqref{eq:vol_drive}) is at $z_0 = 2H$ with width
$\sigma = 0.078H$, sufficiently far from the lower damping zone and permitting
sufficient room for the upward propagating wave to grow as $\propto e^{z/2H}$.
Finally, we want similar grid spacing in the $x$ and $z$ directions (i.e.\ $L_x
/ N_x \sim L_z /N_z$), guided by the intuition that turbulence generated by wave
breaking is approximately isotropic, so we use $L_x = 4H$ and $N_z / N_x = 4$.

@@

After measuring $z_c$ (see Section~\ref{ss:cl_prop}) and $F(z)$
(Eq.~\eqref{eq:F_def}) at each time step, we determine each of $F_i$, $F_a$,
$F_r$, $F_s$ as follows:
\begin{align}
    F_i(t) &= F_{al}A_i^2(t),\\
    F_r(t) &= F_i(t) - \frac{1}{H}
        \int\limits_{z_c - \Delta z - H}^{z_c - \Delta z}F(z, t)\;\mathrm{d}z
        % \ev{F(z, t)}_{z \in [z_c - \Delta z - H, z_c - \Delta z]}
        ,\label{eq:fr_def}\\
    F_s(t) &=
        \frac{1}{\Delta z}\int\limits_{z_c}^{z_c + \Delta z}
            \min(F(z, t), 0)\;\mathrm{d}z,
        % \ev{\z{F(z, t): F(z, t) < 0}}_{z \in [z_c, z_c + \Delta z]},
        \label{eq:fs_def}\\
    F_a(t) &= F_i - F_r - F_s.\label{eq:fa_def}
\end{align}
Fig.~\ref{fig:nl_f_amps2} depicts the four components of this flux
decomposition.
% Here, $\ev{\dots}_{z \in [z_a, z_b]}$ denotes a vertical average over the
% interval $[z_a, z_b]$.
Below the critical layer, we average over an interval of length $H$, also the
vertical wavelength. The offset $\Delta z$ is necessary to make the measurement
of the incident flux unaffected by the turbulence within the critical layer
itself. The width of the critical layer is limited by $\mathrm{Ri} \lesssim 1$
(see Section~\ref{ss:khi}), which bounds its vertical extent $\sim
\frac{1}{\abs{k_{z}}}$. We empirically found an offset of $\Delta z =
\frac{3}{\abs{k_z}}$ was necessary to be sufficiently far from strong
fluctuations near the critical layer.
